Achieving maximum possible precision on stellar radial-velocity measurements in multi-epoch spectroscopy

### 2017-12-04 paper/paper.tex
minor edits

% units macros
\newcommand{\unit}[1]{\mathrm{#1}}
\newcommand{\km}{\unit{km}}
\newcommand{\m}{\unit{m}}
\newcommand{\s}{\unit{s}}
\newcommand{\kms}{\km\,\s^{-1}}
\newcommand{\ms}{\m\,\s^{-1}}

@@

While the general idea of measuring a Doppler shift by the relative motions of stellar absorption lines is fairly straightforward, there are considerable subtleties to implementing this, from the choice of reference spectrum or template to the treatment of telluric absorption features. Ultra-stabilized spectrographs like HARPS \todo{(define \& cite)} predominantly utilize a cross-correlation method with a custom-built list of spectral lines (see e.g. Queloz1995, Baranne1996). 
A lesser-used alternative is cross-correlation or maximum-likelihood fitting using a template spectrum generated from the data (Buchhave, Anglada-Escude). 
This often involves masking or downweighting of the most telluric-contaminated spectral regions.

@@

It is important to state at the outset the assumptions under which are operating for these experiments. 
We assume that our data set consists of multiple observations of the same star with good coverage throughout the observing season, so that we have seen the star many times spanning the range of telluric-to-star relative RV shifts. 
This means we are not considering the case of a real-time data reduction pipeline. 
We also assume a HARPS-like setup, i.e. no absorption cell, but the points we make can be easily generalized to the gas-cell case. 
We return to this point in the discussion \todo{(do this)}.

We additionally assume that the data are extremely well wavelength-calibrated. 
In other work, we will discuss this, but our current belief is that current RV spectrographs are supremely well wavelength-calibrated.

[any other assumptions?]

We begin in Section \ref{s:info} with a derivation of the information-theory-based limit on the RV information content of a spectrum. 
In Section \ref{s:methods}, we summarize common RV extraction practices used by current \RV\ pipelines, including cross-correlation and template fitting. 
We then implement these techniques to extract \RV s using 5-\r{A} section of simulated and real spectra, as described in Section \ref{s:data}. 
We compare the relative performances of the methods in Section \ref{s:experiments} and conclude with some recommendations for future RV pipelines in Section \ref{s:summary}.

@@

The basic model that we use to generate fake spectra is extremely simple, assuming a perfectly normalized continuum; isolated, perfectly Gaussian spectral lines; and white noise only. 
We simulated a 5-$\AA$ region with random noise corresponding to an SNR of 100, a typical value for a single \RV\ spectrum. 
Lines were inserted as Gaussians with prescribed centers, widths, and depths. 
Every line was convolved with the same line spread function, which was taken to be a Gaussian with $\sigma$ = 0.05 $\AA$. 
In the simplest test case, an arbitrary single line at 5000 $\AA$ was used (see Section \ref{s:singleline}). 
However, for most of our tests we injected a set of 12 lines with centers between 4998 - 5002 $\AA$. 
The Gaussian parameters for these lines were based on fits to a reflected Solar spectrum from Vesta taken by \HARPS. \todo{(find citation/program number)} 
The resulting spectrum strongly resembles a 5-$\AA$ chunk taken from a typical \HARPS\ spectrum of a Sun-like star (Figure \ref{fig:spectra}).

@@

\subsection{HARPS Data}

@@

\subsection{HARPS Data}

all the data-driven code works on real data. we get a better RMS on the star from the data-driven template than from the binary mask. based on this, we believe that the HARPS pipeline doesn't strictly use a binary mask.

@@

\acknowledgements
MPIA hospitality, Montet, Foreman-Mackey, Bean, Stuermer, Rix, HARPS team, ESO, Stars group meeting, Brewer

>>>>

% units macros
\newcommand{\unit}[1]{\mathrm{#1}}
\newcommand{\km}{\unit{km}}
\newcommand{\m}{\unit{m}}
\newcommand{\s}{\unit{s}}
\newcommand{\kms}{\km\,\s^{-1}}
\newcommand{\ms}{\m\,\s^{-1}}
\newcommand{\ang}{\text{\normalfont\AA}}

@@

While the general idea of measuring a Doppler shift by the relative motions of stellar absorption lines is fairly straightforward, there are considerable subtleties to implementing this, from the choice of reference spectrum or template to the treatment of telluric absorption features. Ultra-stabilized spectrographs like \HARPS\ \todo{(define \& cite)} predominantly utilize a cross-correlation method with a custom-built list of spectral lines (see e.g. Queloz1995, Baranne1996). 
A lesser-used alternative is cross-correlation or maximum-likelihood fitting using a template spectrum generated from the data (Buchhave, Anglada-Escude). 
This often involves masking or downweighting of the most telluric-contaminated spectral regions.

@@

For the purposes of this work, we assume that our data set consists of multiple observations of the same star with good coverage throughout the observing season, so that we have seen the star many times spanning the range of telluric-to-star relative RV shifts. 
This means we are not considering the case of a real-time data reduction pipeline. 
We also assume a HARPS-like setup, i.e. no absorption cell, but the points we make can be easily generalized to the gas-cell case. 
We return to this point in the discussion \todo{(do this)}.

%wavelength calibration assumption removed b/c it fits better in Section 2... but maybe we want to add it back?

We begin in Section \ref{s:info} with a derivation of the information-theory-based limit on the RV information content of a spectrum. 
In Section \ref{s:methods}, we summarize common RV extraction practices used by current \RV\ pipelines, including cross-correlation and template fitting. 
We then implement these techniques to extract \RV s using 5-\ang-wide sections of simulated and real spectra, as described in Section \ref{s:data}. 
We compare the relative performances of the methods in Section \ref{s:experiments} and conclude with some recommendations for future RV pipelines in Section \ref{s:summary}.

@@

The basic model that we use to generate fake spectra is extremely simple, assuming a perfectly normalized continuum; isolated, perfectly Gaussian spectral lines; and white noise only. 
We simulated a 5-\ang\ region with random noise corresponding to an SNR of 100, a typical value for a single \RV\ spectrum. 
Lines were inserted as Gaussians with prescribed centers, widths, and depths. 
Every line was convolved with the same line spread function, which was taken to be a Gaussian with $\sigma$ = 0.05 \ang. 
In the simplest test case, an arbitrary single line at 5000 \ang\ was used (see Section \ref{s:singleline}). 
However, for most of our tests we injected a set of 12 lines with centers between 4998 - 5002 \ang. 
The Gaussian parameters for these lines were based on fits to a reflected Solar spectrum from Vesta taken by \HARPS. \todo{(find citation/program number)} 
The resulting spectrum strongly resembles a 5-\ang\ chunk taken from a typical \HARPS\ spectrum of a Sun-like star (Figure \ref{fig:spectra}).

@@

\begin{figure}
\centering
% \includegraphics[width=\columnwidth]{spectra}
\caption{}
\label{fig:spectra}
\end{figure}

\subsection{\HARPS\ Data}

@@

\subsection{Real Data}

all the data-driven code works on real data. we get a better RMS on the star from the data-driven template than from the binary mask. based on this, we believe that the \HARPS\ pipeline doesn't strictly use a binary mask.

@@

\acknowledgements
MPIA hospitality, Montet, Foreman-Mackey, Bean, Stuermer, Rix, \HARPS\ team, ESO, Stars group meeting, Brewer
